Supplementary Material for ``Evaluating infectious disease forecasts with allocation scoring rules''

### 2023-11-16 alloscore_manuscript/alloscore-application.tex

The weighted interval score (WIS), is a scoring rule first defined in 2020 to assist with scoring public forecasts that were being made in the early stages of the COVID-19 pandemic.\citep{bracher2021evaluating}
Many COVID forecasts were stored in a quantile format, where probabilistic distributions are encoded by a fixed set of quantiles.
%% optional detail
This was largely a choice of convenience (e.g., all forecasts had a known, fixed data size) and was codified by the US COVID-19 Forecast Hub and other collaborative forecasting projects.\citep{cramer_united_2022}
%%
While pointing a reader interested in more mathematical detail to \cite{bracher2021evaluating}, we note simply that the WIS, is as its name suggests, a weighted sum of interval scores at different probability levels (e.g., 50\% prediction intervals, 80\% PIs, 95\% PIs, etc...).
Larger interval scores indicate less skillful forecasts.
Interval scores themselves consist of (a) the width of the interval, with larger intervals receiving higher scores, and (b) a penalty if the interval does not cover the eventual observation, which increases the further away the interval is from the observed value.
We note that the most commonly used version of WIS is one that chooses a specific set of weights so that WIS approximates the continuous rank probability score (CRPS) or pinball loss, a commonly used score for probabilistic forecasts.
Other specifications of weights for the various interval scores can be used.
\emph{It is important to note that the version of WIS that has been commonly used for evaluating COVID forecasts was proposed because it approximates a commonly used forecast scoring rule, the CRPS, and not because it aligned with any particular public health decision-making rationale.}

The decision-making setup that leads to a choice of WIS as a scoring rule can be formulated using the same recipe as shown in section \ref{sec:methods.detailed.decisiontheory}. In fields such as meteorology and supply chain management, a great deal of attention has been given to the problem where a decision must be made about the quantity of a resource to purchase for a single location in the face of a fixed cost $C$ for each unit of the resource and a loss $L$ that will be incurred for each unit of unmet need.\ngr{I might suggest adding a detailed Step 1/2/3 as an appendix and referencing it here. But I think that adding that full detail here might be too much.} This decision-making problem leads to the pinball loss that is often used in quantile regression, and thence to the CRPS (or WIS, via approximation) as a scoring rule.\ngr{can we make more explicit the link here to the section above? I'm not confident in my ability to describe with technical accuracy the link between quantile/pinball loss and CRPS, but I think this is a good opportunity to do so briefly.}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in gzfile(file, "{}rb"{}): cannot open the connection}}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in group\_by(Kat15k\_alloscores, model, reference\_date): object 'Kat15k\_alloscores' not found}}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in unique(alloscore\_data\_summaries\$model): object 'alloscore\_data\_summaries' not found}}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in `filter()`:\\\#\# i In argument: `model \%in\% alloscore\_models`.\\\#\# Caused by error in `model \%in\% alloscore\_models`:\\\#\# ! object 'alloscore\_models' not found}}\end{kframe}
\end{knitrout}

\subsubsection{Forecast scores showed differences over time}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in `filter()`:\\\#\# i In argument: `target\_end\_date \%in\%\\\#\# \ \ score\_data\_summaries\$target\_end\_date`.\\\#\# Caused by error in `target\_end\_date \%in\% score\_data\_summaries\$target\_end\_date`:\\\#\# ! object 'score\_data\_summaries' not found}}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in ggplot(dat\_to\_plot, aes(x = target\_end\_date, y = value)): object 'dat\_to\_plot' not found}}\end{kframe}
\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in ggplot(alloscore\_data\_summaries, aes(x = target\_end\_date, y = alloscore, : object 'alloscore\_data\_summaries' not found}}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in ggplot(score\_data\_summaries, aes(x = target\_end\_date, y = mwis, : object 'score\_data\_summaries' not found}}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in cowplot::plot\_grid(p1, p2, p4, nrow = 3, align = "{}v"{}, rel\_heights = c(1, : object 'p1' not found}}\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in left\_join(alloscore\_data\_summaries, score\_data\_summaries, by = c("{}model"{}, : object 'alloscore\_data\_summaries' not found}}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in filter(all\_data, as.Date(reference\_date.x) < as.Date("{}2022-02-01"{})): object 'all\_data' not found}}\end{kframe}
\end{knitrout}

>>>>

The weighted interval score (WIS) was proposed in 2020 as a way to score forecasts that were being made in the early stages of the COVID-19 pandemic \citep{bracher2021evaluating}; equivalent scores had also been used in previous forecast evaluation efforts \cite[e.g.,][]{hong2016probabilisticEnergyForecasting}.
The WIS is a proper scoring rule for forecasts that use a set of quantiles to represent a probabilistic forecast distribution.
%% optional detail
Many early COVID-19 modeling efforts, including the US COVID-19 Forecast Hub and other collaborative forecasting projects, adopted a quantile forecast format as a matter of convenience (e.g., this format does not require the forecaster to pre-specify an upper bound for disease counts as would be required to collect forecasts in the format of bin probabilities) \citep{cramer_united_2022}.
%%
While pointing a reader interested in more mathematical detail to \cite{bracher2021evaluating}, we note simply that the WIS is a weighted sum of interval scores at different probability levels (e.g., 50\% prediction intervals, 80\% PIs, 95\% PIs, etc...).
Larger interval scores indicate less skillful forecasts.
An interval score consists of (a) the width of the interval, with larger intervals receiving higher scores, and (b) a penalty if the interval does not cover the eventual observation, which increases the further away the interval is from the observed value.
Equivalently, the WIS can also be characterized as a weighted sum of quantile scores for each individual predictive quantile.
The quantile score for a particular quantile level assigns an asymmetric penalty to predictions that are too high or too low, with the relative sizes of the penalties set so that in expectation the score is minimized by the given quantile of the distribution.
The most commonly used version of WIS is one that uses an equal weighting of all quantile levels, in which case WIS approximates the continuous ranked probability score (CRPS), a commonly used score for probabilistic forecasts.
\apg{Maybe a rewording and/or a supplement reference here to mention that any WIS approximates the a corresponding weighted CRPS, and that WIS actually is the wCRPS for a discrete measure/weighting.}
\emph{It is important to note that this weighting was proposed because the resulting score approximates the CRPS, and not because it aligned with any particular public health decision-making rationale.}

That said, the quantile score and WIS can be derived using the same decision theoretic procedure that we outlined in section \ref{sec:methods.detailed}.
In fields such as meteorology and supply chain management, a great deal of attention has been given to the problem where a decision must be made about the quantity of a resource to purchase for a single location in the face of a fixed cost $C$ for each unit of the resource and a loss $L$ that will be incurred for each unit of unmet need.
This leads to the quantile score for the probability level $\tau = 1 - C/L$.
From this point, the WIS or CRPS can be obtained by averaging across a range of decision making settings with different cost and loss parameters, using a similar motivation that we used to obtain the IAS from the AS in section \ref{sec:methods.detailed.integrated_allocation}.
\apg{Refs to supp for quantile score derivation and detail about averaging.}
%\ngr{I might suggest adding a detailed Step 1/2/3 as an appendix and referencing it here. But I think that adding that full detail here might be too much.}
%\ngr{can we make more explicit the link here to the section above? I'm not confident in my ability to describe with technical accuracy the link between quantile/pinball loss and CRPS, but I think this is a good opportunity to do so briefly.}

@@

\subsubsection{Forecast scores showed differences over time}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/metrics-over-time-1} \caption[Hospital admissions and evaluation metrics over time]{Hospital admissions and evaluation metrics over time. Panel A shows the number of hospital admissions in the US as a whole due to COVID-19 on a sequence of 15 Mondays from December 2021 through March 2022. These are the days for which forecasts were made and evaluated. A horizontal dashed line at 15,000 shows the assumed resource constraint $K$. Panel B shows allocation scores for each model's 14-day ahead forecast, across all US states. The x-axis corresponds to the date for which the prediction was made. Allocation scores typically are high when the observed value is near to the constraint, which occurs during the last Monday in December (on the way up) and the last Monday in January (on the way down). Panel C shows the MWIS metric across weeks, averaged across all states. Similarly to panel B, the x-axis corresponds to the date for which the prediction was made. MWIS values tend to scale with the observed and predicted values, and the peak MWIS values happen around and just after the pean of the Omicron wave.}\label{fig:metrics-over-time}
\end{figure}

\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/metrics-correlation-1} \caption[Association of standardized ranks for MWIS and allocation score by model and week]{Association of standardized ranks for MWIS and allocation score by model and week. Each facet of the plot corresponds to one model. Within each facet, each point corresponds to a week. The x- and y-values correspond to the MWIS standardized rank and the allocation score standardized rank for that week. Points corresponding to earlier dates have darker shading. The size of the point corresponds to the observed value on the date for which the prediction was made. Models show different degrees of association between the two metrics.}\label{fig:metrics-correlation}
\end{figure}

\end{knitrout}

### 2023-11-16 alloscore_manuscript/supplement.rnw

Expanding on the decision theoretic perspective sketched in Section \ref{sec:quantiles_shortage}, we can view a loss function 
as a general tool for formalizing a decision problem that assigns numerical value to the \emph{result} of taking an \emph{action} $x$
in preparation for an \emph{outcome} $y$.  A \emph{scoring rule} $S$ (or just \emph{score}) is a loss function where the action 
is a probabilistic forecast $F$ of the outcome $y$ (or the stating by a forecaster of the forecast), written as $S(F,y)$.

A scoring rule $S$ is \emph{proper} when it    the expected score $\Ex [S(F,Y)]$ is lowest 
 invites the mistaken impression that whether a score 
is proper might depend on the true distribution of what is being forecasted.

>>>>

Expanding on the decision theoretic perspective sketched in Section \ref{sec:quantiles_shortage}, we can view a loss function 
as a general tool for formalizing a decision problem that assigns numerical value to the \emph{result} of taking an \emph{action} $x$
in preparation for an \emph{outcome} $y$.  A \emph{scoring rule} $S$ is a loss function where the action 
is a probabilistic forecast $F$ of the outcome $y$ (or the statement of $F$ by a forecaster). Just as in 
Section \ref{sec:quantiles_shortage}, given an action $F$, $S$ transforms a random outcome variable $Y$ into a random loss $S(F,Y)$. 
We refer to the realized loss $S(F,y)$ as the \emph{score} of $F$ at $y$, and the process of evaluating $S(F,y_i)$ for
a data set $\mathcal{Y}=\{y_i\}$ as \emph{scoring} $F$ against $\mathcal{Y}$.

Decision theoretically, probabilistic forecasts are a unique kind of action in that they can be used to generate their own 
(simulated) outcome data, against which they can be scored using $S$. $S$ therefore commits a probabilistic forecast $F$
to the ``self-assessment'' $\Ex [S(F, Y^F)]$, where $Y^F \sim F$ is the random variable defined by sampling from $F$, as well
to an assessment $\Ex [S(G, Y^F)]$ of any alternative forecast $G$.
A natural consistency criterion for $S$ is that it does not commit $F$ to assessing any other 
forecast $G$ as being better than $F$ itself, that is, that
\begin{align}
\Ex [S(F, Y^F)] \leq \Ex [S(G, Y^F)]
\end{align} 
for any $F,G$. Otherwise, the optimal decision for some forecaster would be to state a forecast $G$ other than the forecast $F$
which they believe describes the outcome $Y$.
A scoring rule meeting this criterion is called \emph{proper}.  The inequality can also be written 
as $\Ex_F [S(F, Y)] \leq \Ex_F [S(G, Y)]$ where the subscript specifies the distribution of $Y$.  
$S$ is \emph{strictly proper} when this inequality is sharp, in which case the \emph{only} optimal decision for a forecaster is to state
the forecast they believe to be true.

Another definition of $S$ being proper that is often used is that $\Ex[S(F,Y)]$ is lowest when $F$ is the true distribution of $Y$.  
Under a flexible reading, this definition is equivalent to ours, but we find it problematic because it invites the mistaken impression that 
the properness of a score might depend on the true distribution of what is being forecasted. Whether a score is proper is unrelated 
to any particular forecast being scored or source of the data being used to score it.

The condition of being proper is quite strong, and na\"{i}ve means of reducing a forecast distribution to a single
number based on an observed outcome $y$ will generally define an improper scoring rule.  For example, the \emph{probability score} 
$S_{\mathrm{Prob},c}(F,y):= -(F(y+c) - F(y-c))$ and the \emph{linear score} $S_{\mathrm{Lin}}
(F,y):= \lim_{c\to 0} \frac{1}{2c}S_{\mathrm{Prob},c}(F,y) = -f(y)$ (where we assume $F$ has a density $f$) and are
both improper because they commit a general $F$ to assessing an alternative forecast $G$ as being better than $F$
itself whenever $G$ is sufficiently more concentrated than $F$ in the neighborhood $[m_f-c,m_f+c]$ of a mode $m_f$ of
$f$. The same is true when the outcome is discrete (and ordered for $S_{\mathrm{Prob}}$) with index $y$: for a forecast
pmf $p(y)$, moving the mass of $p$ onto a neigborhood of a mode $m_p$ to get a new forecast $p_m$ will improve the 
expected score $\Ex[S_{\mathrm{Prob/Lin}}(p_m,Y^p)]$ according to $p$ itself. This creates the classic pathology
of probabilistic forecasters having no incentive to express uncertainty. The meteorologist getting paid according to
$p_{\text{rain}}\text{rain} + (1-p_{\text{rain}})(1-\text{rain})$ (the negative of the linear score) will say rain is 
inevitable or impossible when the chances of rain appear to be 51\% or 49\%.

Sometimes a scoring rule can be made proper with an added correction that penalizes

But given an auxiliary decision problem $\mathcal{D}_{\mathrm{Aux}} =\{\mathcal{X}, \mathcal{Y}, l(x,y)\}$, such as the
allocation problem from the main text or the cost-loss problem from Section \ref{sec:quantiles_shortage}, we can
produce a scoring rule $S_l$ that is automatically proper using the Bayes act formalism introduced in Section \ref
{sec:methods.detailed.specific_allocation}. This works by defining $S_l(F, y)$ as the value $l(x^F,y)$ already given by  
$\mathcal{D}_{\mathrm{Aux}}$ for the Bayes act $x^F$; that is, the action which \emph{by design} will be assessed via
the forecast $F$ to have an expected loss $\Ex[l(x^F,Y^F)]$ as low as possible.

by mapping a forecast $F$ to an the action $x^F$
